Taylor \& Francis \LaTeX\ template for authors (\textsf{Interact} layout + NLM reference style)

### 2021-12-07 paper/interactnlmsample.tex
added ROM equation

>>>>

\section{ROM}
\subsection{ROM Implementation}

Using the POD-Galerkin projection, a ROM was constructed to approximate the group flux and the core power of the LRA benchmark.
As discussed in the previous chapter, the time-dependent diffusion equation coupled with the precursors group equation can be cast into a compact form as in Eq.~\ref{eq:transient fom}.
Similar to the TWIGL benchmark ROM, POD basis set was generated for each group flux while one basis set was generated for both precursors groups, i.e, 
$\mathbf{\Phi}_1(t) = \mathbf{U}_1\mathbf{a}_1(t)$, $\mathbf{\Phi}(t)_2 = \mathbf{U}_2\mathbf{a}_2(t)$ and $\mathbf{C}(t) = \mathbf{U}_c\mathbf{a}_c(t)$, where the $\mathbf{U}_1$, $\mathbf{U}_2$  and $\mathbf{U}_c$ are the POD bases of the fast group, thermal group and precursor concentration respectively, and $\mathbf{a}_1$, $\mathbf{a}_2$ and $\mathbf{a}_c$ are their respective temporal coefficients. 
The first 100 normalized singular values \footnote{The normalization is done be dividing each singular value by the sum of all singular values.} of the group flux are shown in Fig.~\ref{fig:lra singular values}.
\begin{figure}[h!]
	\includegraphics[width=1.0\linewidth]{figures/LRA/LRA_singular_values.pdf}
	\caption{flux normalized singular value.}
	\label{fig:lra singular values}
\end{figure}
Since the focus of this application is on treating the non-linearity, rather than trying different ranks of the POD basis, rank 30 was chosen for each group flux and for the precursor concentration.
Also, the temperature field was approximated by projecting it onto a lower dimensional POD space
\begin{equation}
\mathbf{T}(t) = \mathbf{U}_{\text{temp}}\mathbf{a}_T(t) \, ,
\label{eq:temp pod}
\end{equation}
where $\mathbf{U}_{\text{temp}}\in\mathbb{R}^{n \times r_T} $ is the POD subspace of rank $r_T$ and $\mathbf{a}_T\in \mathbb{R}^{r_T}$ is a vector of the corresponding temporal coefficients. 
To obtain the subspace, snapshots of the temperature were collected at different times for which the SVD was computed.
Shown in Fig.~\ref{fig:lra T singular values} are the first 100 normalized singular values, based on which a POD of rank $10$ was selected for approximating the temperature.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\linewidth]{figures/LRA/LRA_T_singular_values.pdf}
	\caption{temperature normalized singular value.}
	\label{fig:lra T singular values}
\end{figure}
Inserting Eq.~\ref{eq:temp pod} into Eq.~\ref{eq:heatup} and multiplying on the left by $\mathbf{U}_T^T$ \footnote{The superscript T denotes the matrix transpose.} yields
\begin{equation}
\frac{d\mathbf{a}_T(t)}{dt}  = \alpha [\underbrace{\mathbf{U}_{\text{temp}}^T \boldsymbol{\Sigma}_{f1}\mathbf{U}_1}_{\mathbf{U}_{T1}}\mathbf{a_1}(t) + \underbrace{\mathbf{U}_{\text{temp}}^T\boldsymbol{\Sigma}_{f2}\mathbf{U}_2}_{\mathbf{U}_{T2}}\mathbf{a_2}(t)] \, .
\label{eq:reduced temp}
\end{equation}
Note that $\mathbf{U}_{T1}$ and $\mathbf{U}_{T2}$ are time-independent, so they can be precomputed once and stored.
Thus, at each time step the flux temporal coefficients (i.e, $\mathbf{a_1}(t)$ and $\mathbf{a_2}(t)$) are computed by solving the reduced system and then used to compute the temperature coefficients by employing a first-order, backward-difference scheme to Eq.~\ref{eq:reduced temp}.
Similar to the full-order model, a fixed point iteration is used to treat the flux non-linearity.
However, it was observed that the reduced model needed more iterations to converge which could be due to the small value of some of the flux coefficients, hence a convergence criteria was employed based on the relative $\text{L}_2$ norm of the error of flux coefficients between two consecutive iterations 
\begin{equation}
\frac{\| \mathbf{a}^{i-1} - \mathbf{a}^{i}\|}{\|\mathbf{a}_i \|} \le 1\times 10^{-6}  \, , 
\end{equation}
Note that the fast absorption cross section needs to be updated at each time step using Eq.~\ref{eq:doppler} in which the core temperature distribution is required.
In order to avoid reconstructing the full order temperature, 
the DEIM was used to approximate the absorption cross section.
For this purpose, snapshots of which were collected in order to generate a POD subspace that is used to approximate its space and to find a set of interpolation indices as illustrated in Algorithm~\ref{alg:DEIM}.
DEIM of order $10$ was used and hence, the cross section was reconstructed at each time step at only 10 points in the spatial space.
Moreover, since the reactivity insertion was simulated by changing the absorption cross section, the operator $\mathbf{A}$ in Eq.~\ref{Eq:transient rom} exhibits time dependence, 
The MDEIM was used to decompose the operator and hence avoid performing the expensive projection at each time step.
Shown in Fig.~\ref{fig:lra L singular values} are the singular values resulting from the SVD of the serialized operator snapshots. 
Based on the shown plot, a DEIM of order $15$ was used.
\begin{figure}[H]
	\includegraphics[width=1.0\linewidth]{figures/LRA/LRA_L_singular_values.pdf}
	\caption{Operator normalized singular value.}
	\label{fig:lra L singular values}
\end{figure}

%------------------------------------------------------------------------%
\subsection{Results}

\subsubsection{ROM results}
ROMs were constructed for FOMs models with increasing the spatial fidelity, i.e, spatial grids of $22\times 22$, $44\times44$ and $55 \times 55$.
For brevity, we show the POD modes with their temporal coefficients of only the $55 \times 55$  grid in Fig.~\ref{fig:lra flux modes}.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=1.0\linewidth]{../figures/LRA_fast_modes.pdf}
		\caption{Fast flux modes.}
		\label{fig:ss fast flux}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=1.0\linewidth]{../figures/LRA_thermal_modes.pdf}
		\caption{Thermal flux modes.}
		\label{fig:ss thermal flux}
	\end{subfigure}
	\caption{Flux POD modes and their temporal coefficient.}
	\label{fig:lra flux modes}
\end{figure}
Using ROMs with the specifications given in the previous section, the absolute relative error in the predicted powers are computed as shown in Fig.~\ref{fig:deim power error}.
\begin{figure}[H]
	\includegraphics[width=1.0\linewidth]{../figures/LRA_power_deim_relative_error.pdf}
	\caption{Relative error of the power predicted by MDEIM ROM}
	\label{fig:deim power error}
\end{figure}
The maximum flux errors for both groups are shown in Table \ref{tab:lra flux errors}.

\begin{table*}[H]
	\centering
	\begin{tabular}{c|c|c}  
		grid &	fast flux  &  thermal flux \\
		\hline
		$22\times 22$ &    &  1     \\
		$44\times 44$ &    &       \\
		$55\times 55$ &    &       \\
		\hline
	\end{tabular}
	\caption{\bf maximum flux error of the fast and thermal flux (\%)}
	\label{tab:lra flux errors}
\end{table*}
It was interesting to compare the power error with that of a ROM without MDEIM, i.e, with performing the operator projection at each time step.
The power error is shown in Fig.~\ref{fig:rom power error} and apparently, the accuracy is comparable to that in Fig.~\ref{fig:deim power error}, implying an accurate approximation of the problem operator via the MDEIM. 
\begin{figure}[H]
	\includegraphics[width=1.0\linewidth]{../figures/LRA_power_rom_relative_error.pdf}
	\caption{Relative error of the power predicted by ROM.}
	\label{fig:rom power error}
\end{figure}
Also, the spatial-averaged relative error in the approximated temperature was computed and is shown in Fig.~\ref{fig:deim tempearture error}.
\begin{figure}[H]
	\includegraphics[width=1.0\linewidth]{../figures/temperature_relative_error.pdf}
	\caption{Relative error of the temperature predicted by ROM.}
	\label{fig:deim tempearture error}
\end{figure}

As mentioned earlier, the non-linearity of the fast absorption cross section was approximated using DEIM of rank 10.
Shown in Fig.~\ref{fig:cross-section deim indices} are the 10 selected interpolation points on the absorption cross section map at the last time step. i.e, $t=3$ s.
Thus, the temperature is constructed only at these locations from which the cross section is evaluated using Eq.~\ref{eq:doppler}.

\begin{figure}[H]
	\includegraphics[width=1.0\linewidth]{../figures/LRA_XS_interpolation_points_XS_map.pdf}
	\caption{The fast cross section at t=3 with the selected interpolation indices in red color.}
	\label{fig:cross section deim indices}
\end{figure}

The power assembly relative errors are computed and shown in Fig.~\ref{fig:assembly power error fm=5} for the $55\times 55$ grid.
The same errors for the other two grids are given in appendix?.
\begin{figure}[H]
	\includegraphics[width=0.9\linewidth]{../figures/lra_assemb_fm=5.pdf}
	\caption{Assembley power relative error for the $55\times 55$ grid}
	\label{fig:assembly power error fm=5}
\end{figure}

The computational time for the FOM and the ROMS are given in Table~\ref{table:cpu time}.

\begin{table*}[h!]
	\centering
	\caption{\bf computational time of the FOM and the ROMs}
	\label{table:cpu time}
	\begin{tabular}{c|c|c|c}  
		grid &	FOM (s)   &  ROM (s) & MDEIM-ROM (s)    \\
		\hline
		$22\times 22$ & 426   & 140   & 101   \\
		$44\times 44$ &  1652  &   283 &  106  \\
		$55\times 55$ &  2539  & 398   & 110   \\
		\hline
	\end{tabular}
\end{table*}
